|

### 2023-05-12 paper/00-main.Rmd
some writing on multilevel results

abstract: |
  In American football, a pass rush is an attempt by the defensive team to disrupt the offense and prevent the quarterback (QB) from completing a pass. Existing metrics for assessing pass rush are either discrete-time quantities or based on subjective judgement. Using player tracking data, we propose STRAIN, a novel metric for evaluating pass rushers in the National Football League (NFL) at the continuous-time within-play level. Inspired by the concept of strain rate in materials science, STRAIN is a simple and interpretable means for measuring defensive pressure in football. It is a directly-observed statistic as a function of two features: the distance between the pass rusher and QB, and the rate at which this distance is being reduced. Our metric possesses great predictability of pressure and stability over time. We use a multilevel model to isolate the contribution of every pass rusher in a play to the average STRAIN while controlling for their pass blocker matchup, the QB, and play-level information. We apply our approach to NFL data and present results for the first 8 weeks of the 2021 season. In particular, we provide comparisons of STRAIN for different defensive positions and play outcomes, and rankings of the NFL's best pass rushers according to our metric.

bibliography: bibliography.bib
biblio-style: chicago
output: rticles::asa_article
header-includes:
  # \usepackage{mathptmx}
  # \renewcommand{\arraystretch}{0.8}
---

>>>>

abstract: |
  In American football, a pass rush is an attempt by the defensive team to disrupt the offense and prevent the quarterback (QB) from completing a pass. Existing metrics for assessing pass rush performance are either discrete-time quantities or based on subjective judgement. Using player tracking data, we propose STRAIN, a novel metric for evaluating pass rushers in the National Football League (NFL) at the continuous-time within-play level. Inspired by the concept of strain rate in materials science, STRAIN is a simple and interpretable means for measuring defensive pressure in football. It is a directly-observed statistic as a function of two features: the distance between the pass rusher and QB, and the rate at which this distance is being reduced. Our metric possesses great predictability of pressure and stability over time. We use a multilevel model to isolate the contribution of every pass rusher in a play to the average STRAIN while controlling for their pass blocker matchup, the QB, and play-level information. We apply our approach to NFL data and present results for the first 8 weeks of the 2021 season. In particular, we provide comparisons of STRAIN for different defensive positions and play outcomes, and rankings of the NFL's best pass rushers according to our metric.

bibliography: bibliography.bib
biblio-style: chicago
output: rticles::asa_article
header-includes:
  \usepackage{mathptmx}
  # \renewcommand{\arraystretch}{0.8}
---

### 2023-05-12 paper/04-results.Rmd
some writing on multilevel results

## Positional STRAIN curves

@@

There are a variety of quantities we are interested in from the multilevel model fit mentioned in Section \ref{sec:methods}.
First, Table \ref{tab:fixedeff} displays the results for the fixed effects terms.
[INSERT CONCLUSIONS ABOUT THESE VARIABLES]
Next, Table \ref{tab:icc} displays the intraclass correlation coefficients (ICC) for the three different grouped terms, describing the proportion variance explained between the group terms in comparison to the residual variance.
While the residual variance is unsurprisingly the largest value, between the three group terms, we observe the largest ICC for pass rushers, followed by the QB, and the blocker.
This emphasizes how STRAIN is mostly attributable to pass rushers, but is necessary to adjust for opposition and other factors.

@@

\begin{table}
\caption{Fixed-effects coefficients estimates for the multilevel model for average STRAIN. Note that the reference pass rusher level is defensive end (DE), denoted R:DE; and the reference pass blocker level is tackle (T), denoted B:T. \label{tab:fixedeff}}
\centering
\begin{tabular}{lrrrr}
\hline
& estimate & se & $t$-statistic & $p$-value \\ 
\hline
Intercept & 0.378 & 0.009 & 43.310 & 0.000 \\ 
Length of play & -0.004 & 0.000 & -38.214 & 0.000 \\
Number of blockers & -0.064 & 0.003 & -24.788 & 0.000 \\
$I_{\text{\{R:DT\}}}$ & -0.081 & 0.008 & -10.340 & 0.000 \\
$I_{\text{\{R:interior\}}}$ & 0.058 & 0.011 & 5.332 & 0.000 \\
$I_{\text{\{R:NT\}}}$ & -0.106 & 0.012 & -9.174 & 0.000 \\
$I_{\text{\{R:OLB\}}}$ & 0.062 & 0.008 & 7.695 & 0.000 \\
$I_{\text{\{R:secondary\}}}$ & 0.145 & 0.012 & 12.328 & 0.000 \\
$I_{\text{\{B:C\}}}$ & -0.023 & 0.006 & -3.625 & 0.000 \\
$I_{\text{\{B:G\}}}$ & -0.011 & 0.005 & -2.290 & 0.023 \\ 
$I_{\text{\{B:other\}}}$ & -0.013 & 0.007 & -1.911 & 0.056 \\
\hline
\end{tabular}
\end{table}

>>>>

## Positional STRAIN curves {#sec:positional}

@@

[CHECK THIS, MAKE SURE WORDING IS CORRECT/NOT CONFUSING]

There are a variety of quantities we are interested in from the multilevel model fit mentioned in Section \ref{sec:methods}.
First, Table \ref{tab:fixedeff} displays the results for the fixed effects terms.
It appears that that the average STRAIN decreases by 0.042 (95% CI [-0.044, -0.040]) for every extra second in play length, and by 0.064 (95% CI [-0.069, -0.059]) every additional blocker in the offensive unit, after accounting for other covariates in the model.
In other words, NFL pass rushers tend to generate more pressure in plays with shorter length and when facing fewer number of blockers, which totally makes sense.

We observe statistically significant differences among the pass rush and pass block positions in most cases, while controlling for other variables.
The coefficient estimates for the pass rush position terms reveal that players with positions closer to the line of scrimmage (DT and NT) appear to generate less STRAIN (relative to DE) than those that are further back, which is consistent with our results in Section \ref{sec:positional}.
For blockers, centers are those that absorb the most pressure on average, compared to the baseline level tackles, guards, and other offensive positions.

<!-- [INSERT CONCLUSIONS ABOUT THESE VARIABLES] -->
<!-- [SAY SOMETHING ABOUT POSITION] -->

Next, Table \ref{tab:icc} displays the intraclass correlation coefficients (ICC) for the three different grouped terms, describing the proportion variance explained between the group terms in comparison to the residual variance.
While the residual variance is unsurprisingly the largest value, between the three group terms, we observe the largest ICC for pass rushers, followed by the QB, and the blocker.
This emphasizes how STRAIN is mostly attributable to pass rushers, but is necessary to adjust for opposition and other factors.

@@

\begin{table}
\caption{Fixed-effects coefficient estimates for the multilevel model for average STRAIN. Note that the reference pass rusher level is defensive end (DE), denoted R:DE; and the reference pass blocker level is tackle (T), denoted B:T. \label{tab:fixedeff}}
\centering
\begin{tabular}{lrrrr}
\hline
& estimate & se & $t$-statistic & $p$-value \\ 
\hline
Intercept & 0.378 & 0.009 & 43.310 & 0.000 \\ 
Length of play & -0.042 & 0.001 & -38.214 & 0.000 \\
Number of blockers & -0.064 & 0.003 & -24.788 & 0.000 \\
$I_{\text{\{R:DT\}}}$ & -0.081 & 0.008 & -10.340 & 0.000 \\
$I_{\text{\{R:interior\}}}$ & 0.058 & 0.011 & 5.332 & 0.000 \\
$I_{\text{\{R:NT\}}}$ & -0.106 & 0.012 & -9.174 & 0.000 \\
$I_{\text{\{R:OLB\}}}$ & 0.062 & 0.008 & 7.695 & 0.000 \\
$I_{\text{\{R:secondary\}}}$ & 0.145 & 0.012 & 12.328 & 0.000 \\
$I_{\text{\{B:C\}}}$ & -0.023 & 0.006 & -3.625 & 0.000 \\
$I_{\text{\{B:G\}}}$ & -0.011 & 0.005 & -2.290 & 0.023 \\ 
$I_{\text{\{B:other\}}}$ & -0.013 & 0.007 & -1.911 & 0.056 \\
\hline
\end{tabular}
\end{table}
